Untitled

### 2019-04-26 CreateExcelTemplates.Rmd
Major update to Excel script
This script has been totally changed making it much simpler and using the more reliable ONSP instead of the previous datasets. It also cuts out large amounts of superfluous code.

@@

```{r setup}
setwd( "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode")
source("Setup.R")
```

Postocode look up data found at
http://ons.maps.arcgis.com/home/item.html?id=ef72efd6adf64b11a2228f7b3e95deea

Covered by open government licesence
```{r}

setwd(PostcodeLookups)

PstCdLSOA.raw <- read.csv("PCD11_OA11_LSOA11_MSOA11_LAD11_EW_LU_aligned_v2.csv", stringsAsFactors = FALSE)

templatefile<- "Convert Postcode to LSOA.xlsx"

lapply(unique(PstCdLSOA.raw$LAD1), function(LAD){ 
   PstCdLSOA <- PstCdLSOA.raw[PstCdLSOA.raw$LAD11NM == LAD,]

filename <- paste0(LAD,"LSOA.xlsx")
setwd(basewd)
wb <- loadWorkbook("Convert Postcode to LSOA.xlsx")
#input the LSOA lookup table
writeData(wb, sheet = "LSOAdata", PstCdLSOA, colNames = T)
#add in randomly sampled postcodes as an example
writeData(wb, sheet = "InpuNames and codes for administrative geographyt", sample(PstCdLSOA$PCD7, 6), 
          startCol = 2, startRow = 2,
          colNames = F)
setwd(ExcelLookups)
saveWorkbook(wb,filename ,overwrite = T)
  })

The data on ONS is old, so taking a current list of post codes is necessary.

https://www.freemaptools.com/download/outcode-postcodes/postcode-outcodes.csv

Need to check the lattitude and longitude and merge them with the LSOA they fall inside. to create a new list

The England and wales shape files can be downloaded here
http://geoportal.statistics.gov.uk/datasets/da831f80764346889837c72508f046fa_0

```{r}

setwd(ShapeFiles)
list.files(pattern = ".shp")

#load shape file
shape <- readOGR(dsn = "LSOA_2004_London_Low_Resolution.shp")

#Convert to Latitude and longitude
shape <- spTransform(shape, CRS("+init=epsg:4326"))

#read postcodes
# setwd(basewd)
# postcodes <-read.csv("ukpostcodes.csv")
# 
# #subset to souteast
# p2 <- postcodes[grepl("^SE",postcodes$postcode),]

#Plot using base this doesn't work I don't know why
# coordinates(p2) <- ~ latitude + longitude  # pts needs to be a data.frame for this to work
# proj4string(p2) <- proj4string(shape)
# 
# plot(shape)
# plot(p2, col="red" , add=TRUE)

#convert shape files to data frame
AG <- tidy(shape)

#Plot shape files as map of london with south east plotted over the top. this does work as expected.
ggplot()+ geom_polygon(data=AG, aes(long, lat, group = group), 
colour = alpha("darkred", 1/2), size = 0.7, fill = 'skyblue', alpha = .3) +geom_point(data= p2, aes(x = longitude, y= latitude))+
  ggtitle("London With SE postcodes overplotted")

#Subset map to a smaller area
AG2 <- AG %>% filter(long>-0.2, long<0.1, lat<51.55, lat>51.45)
#check that the points are aligned. They are.
ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group), 
colour = alpha("darkred", 1/2), size = 0.7, fill = 'skyblue', alpha = .3) + geom_point(data= p2, aes(x = longitude, y= latitude))

```{r}
p3 <- postcodes %>% filter(latitude<65)
#ensure that the points are converted to latitude and longitude
coordinates(p3) <- ~longitude + latitude
#set the projection of the data to longitude and latitude
proj4string(p3) <- CRS("+proj=longlat")
#I don't undertsand what is going on here
p3 <- spTransform(p3, proj4string(shape))
proj4string(p3) <- proj4string(shape)
z <- over(p3, shape)

PstCdLSOA.raw <- bind_cols(postcodes %>% filter(latitude<65), z)

PstCdLSOA.raw <- PstCdLSOA.raw %>% filter(!is.na(PstCdLSOA.raw$LSOA_CODE))

head(PstCdLSOA.raw,100)
```

```{r}
lapply(unique(PstCdLSOA.raw$LA_NAME), function(LAD){ 
   PstCdLSOA <- PstCdLSOA.raw[PstCdLSOA.raw$LA_NAME == LAD,]

filename <- paste0(LAD,"LSOA.xlsx")
setwd(basewd)
wb <- loadWorkbook("Convert Postcode to LSOA.xlsx")
#input the LSOA lookup table
writeData(wb, sheet = "LSOAdata", PstCdLSOA, colNames = T)
#add in randomly sampled postcodes as an example
writeData(wb, sheet = "Input", sample(PstCdLSOA$postcode, 6), 
          startCol = 2, startRow = 2,
          colNames = F)
setwd(ExcelLookups)
saveWorkbook(wb,filename ,overwrite = T)
  })
```

```{r}
#Read postcodes from the ONS folder of codes
setwd(ONSpostcodes)
list.files()
p2<- read.csv("se.csv", header = FALSE)
setwd(ONSpostcodeMeta)
names(p2) <- z <- read.csv("Code-Point_Open_Column_Headers.csv")[1,] %>%t

#Check that the shapes are using the same coordinate system
AG <- tidy(shape)
#Subset map to a smaller area
AG2 <- AG %>% filter(long>530000, long<550000, lat>170000, lat<185000)
#check that the points are aligned. They are.
ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group), 
colour = alpha("darkred", 1/2), size = 0.7, fill = 'skyblue', alpha = .3)+ geom_point(data= p2, aes(x = Eastings, y= Northings))

#set the Northings an eastings columns to be the coordinates
coordinates(p2) <- ~Eastings + Northings
#defines the coordinate system
proj4string(p2) <- CRS("+init=epsg:27700")
#converts the coordinate system from the one definded to the coordinate system of the shape file, this step is not neccesary in this case as both are using the Uk grid referencing system.
p2 <- spTransform(p2, proj4string(shape))

#check that the conversion has worked properly
plot(shape)
plot(p2, col="red" , add=TRUE)

#Check which LSOA each postcode falls within and bind it to the postcode data
PstCdLSOA.raw <- bind_cols(tidy(p2), over(p2, shape))

#remove any NA values
PstCdLSOA.raw <- PstCdLSOA.raw %>% filter(!is.na(PstCdLSOA.raw$LSOA_CODE))

This code block works smoothly using government data testing on the "se" postcodes. The next code chunk will perform the same operation for all postcodes. as a result it is computationally intense.

#Using only ONS data

Load the LSOA shape data for EW 
```{r}
setwd(LSOAshapedata)

shape <- readOGR(dsn = list.files(pattern = "shp"))

# Load Post code data

This chunk checks the exiatane of the postcode to lsoa lookup if it doesn't find it, the second part of the code is run that creates the lookup. Creating the lookup takes quite a lot of time so it is good to only have to do it once.
```{r}

if(file.exists(file.path(DataFolder,"PstCdLSOA.rds"))){
 PstCdLSOA.raw <- readRDS(file.path(DataFolder,"PstCdLSOA.rds"))
} else {
  source(file.path(CommonCode, "PostcodeLookup.R"))
}

Add in LAD names and Region, remove non-essential columns
```{r}
setwd(AddGeog)
list.files()

LADReg <- read.csv("LAD15_RGN15_EN_LU.csv", stringsAsFactors = FALSE   )
Wardnames <- read.csv("Ward_to_Local_Authority_District_to_County_to_Region_to_Country_December_2016_Lookup_in_United_Kingdom_V2.csv", stringsAsFactors = FALSE) %>%
  select(WD16CD, WD16NM)

PstCdLSOA <- PstCdLSOA.raw %>% mutate_if(., is.factor, as.character)

#There are a couple of errors is the region LSOA look up these are manually corrected here.
PstCdLSOA[PstCdLSOA$Postcode=="BN994AE", c(3,4,7)] <- PstCdLSOA[PstCdLSOA$Postcode=="BN126BW", c(3,4,7)]

PstCdLSOA[PstCdLSOA$Postcode=="DY4 4BA", c(3,4,7)] <- PstCdLSOA[PstCdLSOA$Postcode=="DY4 0YX", c(3,4,7)] 
PstCdLSOA[PstCdLSOA$Postcode=="DY4 4BB", c(3,4,7)] <- PstCdLSOA[PstCdLSOA$Postcode=="DY4 0YX", c(3,4,7)]

PstCdLSOA <- PstCdLSOA %>% left_join(., LADReg, by = c("Admin_district_code"="LAD15CD")) %>%
  left_join(., Wardnames, c("Admin_ward_code" = "WD16CD")) %>% 
  select(Postcode, Country_code, Admin_district_code, 
                          Admin_ward_code, lsoa11cd, lsoa11nm, LAD15NM, 
                          RGN15CD, RGN15NM, WD16NM)

#Wales doesn't have any LAD for some reason, fix is to add in LSOA name subtracting the code

PstCdLSOA$LAD15NM[grepl("W", PstCdLSOA$Country_code)] <- gsub(' .*', '' , PstCdLSOA$lsoa11nm[grepl("W", PstCdLSOA$Country_code)])

#Make sure that Scotland and Wales are included in the regions.

PstCdLSOA[grepl("W", PstCdLSOA$Country_code),c("RGN15CD" , "RGN15NM" )] <- c("Wales", "Wales")

PstCdLSOA[grepl("S", PstCdLSOA$Country_code),c("RGN15CD" , "RGN15NM" )] <- c("Scotland", "Scotland")

#remove spaces from the the Postcodes

PstCdLSOA <- PstCdLSOA %>% mutate(Postcode = gsub(" ", "", Postcode))

@@

lapply(unique(PstCdLSOA$RGN15NM), function(n){
  title <- paste("Region", n)
  dir.create(file.path(Regions, title ), showWarnings = FALSE)
  #The inner loop creates the excel files for each LAD within each region
  PstCdLSOA.Reg <- PstCdLSOA %>% filter(RGN15NM == n)

  print(title)
  lapply(unique(PstCdLSOA.Reg$LAD15NM), function(LAD){

    PstCdLSOA.filt <- PstCdLSOA.Reg %>% filter(LAD15NM == LAD)
    print(paste(n, ":", LAD))
      ExemptionsFile <- paste0(LAD,"ExemptionsLSOA.xlsx")
      CTBandsFile <- paste0(LAD,"CTBandsLSOA.xlsx")
      setwd(basewd)
      wb <- loadWorkbook("Convert Postcode to LSOA.xlsx")
      #input the LSOA lookup table
      writeData(wb, sheet = "LSOAdata", PstCdLSOA.filt, colNames = T)
      #add in randomly sampled postcodes as an example
      writeData(wb, sheet = "Input", sample(PstCdLSOA.filt$Postcode, 6),
                startCol = 2, startRow = 2,
                colNames = F)
      setwd(file.path(Regions, title))
      saveWorkbook(wb,ExemptionsFile ,overwrite = T)
      saveWorkbook(wb,CTBandsFile ,overwrite = T)
    })
})

>>>>

#references

The key data is the ONS postcode directory. This data is released quarterly, and contains all the postcode past and present.
This data set is great as it matches the postcodes to LSOA which makes the whole geospatial part much easier.

The other two data sets are the regional code look up

and the ward look up. Ward data can be found by searching the ONS open geograph portal. the name of the dataset is similar too "Lower Layer Super Output Area (2011) to Ward (2018) Lookup in England and Wales v3" this dataset is updated regulalry as well the most recent should be used.

This code is greatly simplified from earlier versions due to switching to the ONSPD and removeing the mapping experiments.

@@

```{r}
basewd <-"~/Dropbox/SSE/Empty Homes"
ONSpostcodes <- file.path(basewd, "ONSPD")
AddGeog <- file.path(basewd, "AdditionalGeographies")
wards_file <- "Lower_Layer_Super_Output_Area_2011_to_Ward_2018_Lookup_in_England_and_Wales_v3.csv"
Regions <- file.path(basewd, "Regions")

select <- dplyr::select

Wardnames <- read_csv(file.path(AddGeog, wards_file))

LADReg <- read.csv(file.path( AddGeog, "LAD15_RGN15_EN_LU.csv"), stringsAsFactors = FALSE   )

CorePstCd  <- read_csv(file.path(ONSpostcodes, 
                              "ONSPD_FEB_2019_UK.csv")) %>%
  select(Postcode = pcd, LSOA11CD = lsoa11, 
         MSOA11CD = msoa11, LAD11CD = oslaua, 
         OA11CD = oa11, Country_code = ctry, 
         Region =  rgn)

{
PstCdLSOA <- CorePstCd %>% 
  left_join(., LADReg, by = c("LAD11CD"="LAD15CD")) %>%
  left_join(Wardnames, by = "LSOA11CD") %>%
  #filter(Country_code == "W92000004")
  select(Postcode, Country_code, Admin_district_code = LAD11CD, 
         Admin_ward_code = WD18CD, lsoa11cd = LSOA11CD, LAD15NM = LAD18NM, 
         RGN15CD, RGN15NM, WD16NM = WD18NM) %>%
  mutate(
    #Make sure Scotland and Wales have region names
    RGN15CD = case_when(
      grepl("W", Country_code) ~"Wales",
      grepl("S", Country_code) ~"Scotland",
      TRUE ~ RGN15CD),
    RGN15NM = case_when(
      grepl("W", Country_code) ~"Wales",
      grepl("S", Country_code) ~"Scotland",
      TRUE ~ RGN15NM
    ),
    Postcode = gsub(" ", "", Postcode)) %>% #remove spaces from postcodes
filter(!is.na(RGN15NM),
       LAD15NM != "NA",
       RGN15CD !="Scotland")
#This is a large dataframe so we want to remove it when possible to save memory

#region cleaning
#some LSOA are in multiple regions and Adminwards this code fixes this by using a majority vote

MajRegion <- PstCdLSOA  %>% group_by(lsoa11cd, RGN15CD) %>%
  summarise(count = n(),
            RGN15NM = first(RGN15NM)) %>%
  group_by(lsoa11cd) %>%
  arrange(-count) %>%
  summarise_all(first) %>%
  select(-count)

MajLAD <- PstCdLSOA  %>% group_by(lsoa11cd, Admin_district_code) %>%
  summarise(count = n()) %>%
  group_by(lsoa11cd) %>%
  arrange(-count) %>%
  summarise_all(first) %>%
  select(-count)

PstCdLSOA  <- PstCdLSOA %>%
  select(-Admin_district_code,-RGN15CD, -RGN15NM) %>%
  left_join(MajLAD) %>%
  left_join(MajRegion)
}

rm(CorePstCd)

@@

unique(PstCdLSOA$RGN15NM) %>%
  walk(~{

    title <- paste("Region", .x)
    dir.create(file.path(Regions, title ), showWarnings = FALSE)
    #The inner loop creates the excel files for each LAD within each region
    PstCdLSOA.Reg <- PstCdLSOA %>% filter(RGN15NM == .x)

    print(title)

    unique(PstCdLSOA.Reg$LAD15NM) %>%
      walk(~{

        PstCdLSOA.filt <- PstCdLSOA.Reg %>% filter(LAD15NM == .x)
        print(paste(title, ":", .x))
        ExemptionsFile <- paste0(.x,"ExemptionsLSOA.xlsx")
        DiscountsFile <- paste0(.x,"DiscountsLSOA.xlsx")
        setwd(basewd)
        wb <- loadWorkbook("Convert Postcode to LSOA.xlsx")
        #input the LSOA lookup table
        writeData(wb, sheet = "LSOAdata", PstCdLSOA.filt, colNames = T)
        #add in randomly sampled postcodes as an example
        writeData(wb, sheet = "Input", sample(PstCdLSOA.filt$Postcode, 6),
                  startCol = 2, startRow = 2,
                  colNames = F)
        setwd(file.path(Regions, title))
        saveWorkbook(wb,ExemptionsFile ,overwrite = T)
        writeData(wb, "Input", "Discount_Class") #So that it says Discounts not exemptions, so nobody gets confused.
        saveWorkbook(wb,DiscountsFile ,overwrite = T)

      })

  })
